Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-30 demystify.tex
updates

\section{Is an MS required for admission to PhD in CS in the US?}\label{sec:msrequirement}
\sectioninfo{You do not need an MS to do PhD in CS.}

@@

>>>>

\section{Is MS required for admission to PhD in CS?}\label{sec:msrequirement}
\sectioninfo{You do not need an MS to do PhD in CS.}

@@

\section{``Chance me'' on getting into MIT or Stanford or Ivy League schools}\label{sec:chance-me}

Many students, especially international students, aim for very top CS PhD programs such as Stanford and MIT, and Ivy Leagues. Every year Reddit and other forums have numerious posts asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts), and then later posts on being \emph{``rejected everywhere''}.

\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that you are very likely NOT getting in any of these. Why? people who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be MIT-alum or are academic celebrities who've sent many of their students there. If you are not in this category (doubt you are, because you're asking random people on the internet instead of your trusted LoR writers), then you're likely not going to get in, and MIT and Stanford would glady take your application money.

\paragraph{Top in US News might not really top in CS} Morever, your ``top'' schools might not be top in CS. The Ivies? yes good for many things but usually not coding marathon and typically MIA at top CS confs (ok, they are still good, but a school somewhere in Maryland, Wisconsin, or Urbana Champain cornfield would probably beat them by far). MIT and Stanford? yes, they are top, but many are equivalent, and there could be CS areas where they are not as strong as other schools.  Caltech and John Hopkins?  ... are you confused CS with Physics and Medicine?

You're missing out some serious places: UCSD, UW, UMD, Michigan, Purdue, Stony brook. These all have super strong CS programs but rarely get mentioned. Utah, Rutgers, Buffalo, Penn State, UVA, Ohio, UC Riverside, George Mason ... any of them rings a bell? These might not ranked very high on QS or US News and are not household names to you, your friends, mom and dad, but they are excellent CS schools where graduating students from your dream schools often apply for faculty positions (e.g., see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{GMU stats}). You can totally go to these places and be superstar there.

\paragraph{You don't need to be top}  It is understandable top schools will open doors to more opportunities (you are likely from a very good schools in your country to aim for a PhD in CS in the US). In countries like Vietnam, there are usually only a few top schools, and if you don't get into them, you're doomed (or so you think). But that is not the case in the US. There are simply too many good schools, and it is not a binary choice between top X or nothing.  In fact, ``equity'' is emphasized in the US and many ``smaller'' schools actually are given more resources and attention to help students succeed.

In short, don't aim for just the moon and the stars, also try the sky, clouds, or even just the trees.  You might be happier there and it might catapult you to the sun and beyond.
